Fig2: Candidate genes

### 2020-05-06 analysis/figure-2.Rmd
now with better fig 2, start of fig 3

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
          gene %in% whichgenes,
          sex %in% whichsex,
          treatment != "control") %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    ggplot(aes(x = treatment, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    facet_wrap(~gene, scales = "free_y", nrow = 2) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          strip.text.y = element_text(face = "italic"),
          strip.text.x = element_text(face = "italic"),
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(y = paste(whichsex, "gene expression", sep = " "),
         x = NULL) +
    geom_signif(comparisons = list(c( "bldg", "lay"),
                                 c( "lay", "inc.d3"),
                                 c("inc.d3", "inc.d9"),
                                 c( "inc.d9", "inc.d17"),
                                 c( "inc.d17", "hatch"),
                                 c("hatch", "n5"),
                                 c( "n5", "n9")),  
                map_signif_level=TRUE,
                textsize = 1.5, family = 'Helvetica',
                vjust = 1.5, size = 0)

@@

scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
  p <- ggplot(df, aes(x = gene1, y = gene2)) +
    labs(subtitle = " ", x = myxlab, y = myylab) + 
    scale_color_manual(values = allcolors) +
    geom_smooth(method = "glm",  color = mylinecolor) +
    geom_point(aes(color = treatment))  +
    theme_B3() +  
    theme(axis.title = element_text(face = "italic"), 
            legend.position = "none") +
    stat_cor( size = 2)
  return(p)
}
```

@@

## plot sig candidate genes
a <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "Hypothalamus")
e <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "Pituitary")
i <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = "female")+ labs(title = "Gonad")

b <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
f <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = "male") + labs(title = "")
j <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = "male") + labs(title = "")

c <- candidateboxplot("hypothalamus", c("DRD1", "HTR2C"), "female") + labs(x = NULL, title = "") 
g <- candidateboxplot("pituitary", c( "AVPR1A", "CRHR1" ), "female") + labs(x = NULL, title = "")
k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "")

d <- candidateboxplot("hypothalamus",  c("DRD1", "HTR2C"), "male") + labs(x = NULL, title = " ") 
h <- candidateboxplot("pituitary", c("AVPR1A", "CRHR1"), "male") + labs(x = NULL, title = " ")
l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ")

plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
          rel_widths = c(0.8,0.8,1,1), rel_heights = c(1,1,1.2),
          labels = c("A", " ", "B", "",
                     "C", " ", "D", "",
                     "E", " ", "F", ""), label_size =  8)

@@

## suplement correlations

```{r supplfig2, fig.width = 7.25, fig.height = 7.25}

a <- plotcorrplot(hyp1, "females") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
c <- plotcorrplot(pit1, "females") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
e <- plotcorrplot(gon1, "females") + theme(legend.position = "bottom")  + labs(title = "")

b <- plotcorrplot(hyp2, "males") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
d <- plotcorrplot(pit2, "males") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
f <- plotcorrplot(gon2, "males") + theme(legend.position = "bottom")  + labs(title = "")

plot_grid(a,b,c,d,e,f, nrow = 3, rel_heights = c(1,1,1.5))

>>>>

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
          gene %in% whichgenes,
          sex %in% whichsex) %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    ggplot(aes(x = treatment, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    facet_wrap(~sex, scales = "free_y", nrow = 1) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          axis.title.y = element_text(face = "italic"),
          axis.text.x = element_blank(),
          axis.ticks = element_blank()) +
    labs(y = whichgenes,
         x = NULL) +
    geom_signif(comparisons = list(c( "control", "bldg"),
                                 c( "bldg", "lay"),
                                 c( "lay", "inc.d3"),
                                 c("inc.d3", "inc.d9"),
                                 c( "inc.d9", "inc.d17"),
                                 c( "inc.d17", "hatch"),
                                 c("hatch", "n5"),
                                 c( "n5", "n9")),  
                map_signif_level=TRUE,
                textsize = 1.5, family = 'Helvetica',
                vjust = 1.5, size = 0)

@@

scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
  p <- ggplot(df, aes(x = gene1, y = gene2)) +
    labs(subtitle = " ", x = myxlab, y = myylab) + 
    scale_color_manual(values = allcolors) +
    geom_smooth(method = "glm",  color = mylinecolor) +
    geom_point(aes(color = treatment))  +
    theme_B3() +  
    theme(axis.title = element_text(face = "italic"), 
            legend.position = "none") +
    stat_cor( size = 2)
  return(p)
}

@@

header <- png::readPNG("../figures/images/fig_fig2a.png")
header <- ggdraw() +  draw_image(header, scale = 1)

b <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "Hypothalamus")
g <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(title = "Pituitary")
l <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(title = "Gonad")

c <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "")
h <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL)+ labs(title = "")
m <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)+ labs(title = "")

d <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
i <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) + labs(title = "")
n <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL) + labs(title = "")

d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL , title = " ")  + theme(axis.line.x = element_blank()) 
d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = "Parental stage" ) + theme(strip.text = element_blank())

g1 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL, title = " " )  + theme( strip.text = element_blank(), axis.line.x = element_blank())
g2 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = "Parental stage"  )  + theme(strip.text = element_blank())

j1 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL , title = " ") + theme(strip.text = element_blank(), axis.line.x = element_blank())
j2 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = "Parental stage"  ) + theme(strip.text = element_blank())

btn <- plot_grid(b,c,d, g,h,i,l,m,n, ncol = 3, labels = c("B", "C", " ","E", "F", " ", "H", "I"), label_size = 8, rel_widths = c(1.4,1,1), rel_heights = c(1.1,1,1))
dgj <- plot_grid(d1,d2,g1,g2,j1,j2, ncol = 1, labels = c("D", " ", "G", " " , "J"), label_size = 8, rel_heights = c(1.3,1,1.2,1,1.2,1))
btoj <- plot_grid(btn,dgj, ncol = 2 , rel_widths = c(1.75,1))

plot_grid(header, btoj,  rel_heights = c(0.2,1), ncol = 1)

@@

```{r fig3, fig.height=3, fig.width=7.25}

musicplot <- function(whichgene, whichtissue){

    p <- candidatevsd %>%
      group_by(treatment, tissue, gene, sex)  %>% 
      summarize(median = median(counts, na.rm = T), 
                se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
      dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
      dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
      filter(
        gene %in% whichgene,
        tissue %in% whichtissue      ) %>% 
      collect() %>%
      drop_na() %>%
      ggplot( aes(x = treatment, y = median)) +
      geom_errorbar(aes(ymin = median - se, 
                        ymax = median + se, color = "white"),  width=0) +
      geom_image(aes(image=image), size = 0.1)+
      theme_B3() +
      theme(legend.position = "none",
            title = element_text(face = "italic"),
            strip.text = element_text(color = "white"),
            axis.text.x = element_text(angle = 45, hjust = 1),
            axis.text.y = element_blank(),
            axis.line = element_blank(), axis.ticks = element_blank()) +
      scale_color_manual(values = allcolors) +
      labs(y = " ",x = NULL ) +
      facet_wrap(~sex, nrow = 2, scales = "free_y")
      return(p)
}

music <- png::readPNG("../figures/images/fig_music.png")
music <- ggdraw() +  draw_image(music, scale = 1)

musicalgenes <- png::readPNG("../figures/images/fig_musicalgenes.png")
musicalgenes <- ggdraw() +  draw_image(musicalgenes, scale = 1)

p1 <- musicplot(c("AVP", "OXT"), "hypothalamus") + labs(subtitle = "Hypothalamic AVP & OXT" ) + theme(axis.text.x = element_blank())
p2 <- musicplot(c("AVP", "OXT"), "pituitary") + labs(subtitle = "Pituitary AVP & OXT " ) + theme(axis.text.x = element_blank())
p3 <- musicplot(c("AVP", "OXT"), "gonad") + labs(subtitle = "Gonadal  AVP & OXT"  )

p123 <- plot_grid(music, p1, music, p3, ncol = 2, rel_widths = c(0.125,1))

plot_grid(musicalgenes, p123, nrow = 1, rel_widths = c(1.5,1))
